---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CNAQI

<!-- badges: start -->
<!-- badges: end -->

The goal of chnqir is to calculate Air Quality Index according to regulations and standards of Ministry of Ecology and Environment of China.

## Installation

You can install the development version of chnqir like so:

``` r
devtools::install.package("evanliu3594/cnaqir")
```

## CNAQI

根据中华人民共和国生态环境部《环境空气质量指数（AQI）技术规定（HJ633—2012）》中的相关标准和方法，提供五个用于计算AQI的函数


### `DaliyMeanConc()`
`DaliyMeanConc()` calculates daily mean concentration of 6 pollutants. Note that average `O3` concentration for one day is presented by maximum 8-hour average concentration.

> Note: units of $SO_2$,  $NO_2$,  $PM_{10}$,  $PM_{2.5}$  and  $O_3$ input are  $μg/m^3$

> unit of $CO$ is $mg/m^3$


```{r}
library(CNAQI)

C <- rnorm(24, 35, 5)
DaliyMeanConc("PM2.5", C)
```

### `IAQI_hourly()`
`IAQI_hourly()` is for calculate hourly IAQI for a given pollutant
```{r}
IAQI_hourly("O3", 142)
```

### `AQI_Hourly()`
`AQI_Hourly()` calculates the hourly AQI score and the featured pollutants based on 4 pollutants' concentration.

```{r}
AQI_Hourly(SO2 = 300,NO2 = 200,CO = 100,O3 = 120)
```

### `IAQI_Daily()`
`IAQI_Daily()` calculates daily IAQI for given pollutant

```{r}
IAQI_Daily("O3", 142)
```

### `AQI_Daily()`
`AQI_Daily()` calculate the daily AQI score and the featured pollutants based on 6 pollutants' concentration.

```{r}
AQI_Daily(SO2 = 55, NO2 = 23, CO = 12, O3 = 122, PM2.5 = 35, PM10 = 55)
```

### A sample workflow with AQI series functions in `tidyr` syntax.

#### generating sample data

```{r}
library(tidyverse)

set.seed(3594)

sampleConc <- data.frame(
  Time = 0:23,
  SO2 = rnorm(24, mean = 300, sd = 20),
  NO2 = rnorm(24, mean = 800, sd = 100),
  O3 = rnorm(24, mean = 50, sd = 15),
  CO = rnorm(24, mean = 25, sd = 10),
  PM2.5 = rnorm(24, mean = 200, sd = 50),
  PM10 = rnorm(24, mean = 350, sd = 100)
)

sampleConc
```

#### Calculate hourly IAQI
```{r}
sampleConc %>% 
  pivot_longer(-Time, names_to = 'Pollu', values_to = 'Conc') %>% rowwise() %>% 
  mutate(IAQI_h = IAQI_hourly(Pollu = Pollu, Conc = Conc)) %>% select(-Conc) %>% 
  pivot_wider(names_from = 'Pollu', values_from = 'IAQI_h')

```
#### Calculate Hourly AQI 
```{r}
sampleConc %>% rowwise() %>% mutate(
  AQI_h = AQI_Hourly(SO2 = SO2, NO2 = NO2,CO = CO,O3 = O3), .keep = 'none'
)
```
#### Calculate Daily Concentration and daily AQI
```{r}
sampleConc_daily_mean <- sampleConc %>% select(-Time) %>% 
  imap_dbl(~DaliyMeanConc(Pollu = .y, Conc = .x))

sampleConc_daily_mean
```

#### Calculate daily IAQI
```{r}
sampleConc_daily_mean %>% imap_dbl(~IAQI_Daily(Pollu = .y, Conc = .x))
```

#### Calculate Daily AQI
```{r}
AQI_Daily(SO2 = sampleConc_daily_mean[["SO2"]],
          NO2 = sampleConc_daily_mean[["NO2"]],
          CO = sampleConc_daily_mean[["CO"]],
          PM10 = sampleConc_daily_mean[["PM10"]],
          PM2.5 = sampleConc_daily_mean[["PM2.5"]],
          O3 = sampleConc_daily_mean[["O3"]])
```

